Trajectories

### 2021-09-15 Rmd/report.Rmd
Fix figures

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  echo = TRUE,
  warning = TRUE,
  dev.args = list(png = list(type = "cairo")),
  include = FALSE,
  out.width = "100%",
  results = "asis",
  dpi = 300
)
options(
  knitr.kable.NA = "-",
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
  ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r items_cognate, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
by_cognate <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = cognate, fill = class)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Cognateness", y = "Number of items", fill = "Class")

@@

```{r participants_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
  group_by(time_stamp, lp, dominant_language) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(lp, dominant_language) %>% 
  mutate(n = cumsum(n)) %>% 
  ggplot(aes(time_stamp, n, colour = interaction(lp, dominant_language, sep = " - "), fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Number of responses", colour = "LP - Dominant language") +
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
  )

@@

```{r participants_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r participants_lp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
a <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

@@

$$
\begin{aligned}
\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& (\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& (\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& (\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& (\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& (\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \varepsilon_{pi}
\end{aligned}
$$

@@

```{stan stan_model, echo=TRUE, output.var="model"}

@@

```{r vocabulary_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r vocabulary_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r raw, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  group_by(te, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
  facet_wrap(vars(type)) +
  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(age = as.factor(round(age))) %>% 
  group_by(te, age, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
  facet_wrap(vars(type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

@@

```{r ppc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

@@

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

```{r fix_effets_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

@@

```{r random_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r re_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r emmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r diagnostics_traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  echo = TRUE,
  warning = TRUE,
  dev.args = list(png = list(type = "cairo")),
  include = TRUE,
  out.width = "100%",
  results = "asis",
  dpi = 300
)
options(
  knitr.kable.NA = "-",
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
  ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

```{r items_frequency, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r items_cognate, message=FALSE, warning=FALSE, paged.print=FALSE}
by_cognate <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = cognate, fill = class)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Cognateness", y = "Number of items", fill = "Class")

@@

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
  group_by(time_stamp, lp, dominant_language) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(lp, dominant_language) %>% 
  mutate(n = cumsum(n)) %>% 
  ggplot(aes(time_stamp, n, colour = interaction(lp, dominant_language, sep = " - "), fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Number of responses", colour = "LP - Dominant language") +
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
  )

@@

```{r participants_age, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
a <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

@@

$$
\begin{aligned}
\color{white}
\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& (\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& (\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& (\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& (\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& (\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \varepsilon_{pi}
\end{aligned}
$$

@@

```{stan stan_model, output.var="model"}

@@

```{r vocabulary_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r vocabulary_production, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r vocabulary_long_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  group_by(te, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
  facet_wrap(vars(type)) +
  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(age = as.factor(round(age))) %>% 
  group_by(te, age, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
  facet_wrap(vars(type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

@@

```{r ppc, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

@@

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

```{r fix_effets_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

@@

```{r random_participant, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r re_item, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

```{r diagnostics_traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}
